---
title: "Plasma Analysis on PTSD Project"
author: "YI XIAO"
date: "1/17/2019"
output: rmarkdown::github_document
 
---

```{r}

plasma=read.csv('/Users/xiaoyi/Documents/2018Fall/Clinical/PTSD/plasma.csv',na.strings=c("","NA"))
#?blood pressure have many columns,skip it for now
grep("pressure",colnames(plasma),value = TRUE)
col=c("id","Timepoints","PTSDorNot","Treatment", "Batch_A","Batch_S", "age" , "biol_sex" , "ethnicity_id", "bmi", "education", "marital_status", grep("Mean",colnames(plasma),value = TRUE))
#prepare data
dat1=plasma[,col]
colnames(dat1)
biom=grep("Mean",colnames(plasma),value = TRUE)
options(scipen=10000)
#
#RECODE  the 1,0 into PTSD active 
dat1$Group=ifelse(dat1$PTSDorNot==1,'PTSD','non-PTSD')
dat1$Treatment_new=ifelse(dat1$Treatment==1,'active','sham')
dat1$PTSDorNot=factor(dat1$PTSDorNot)
#col1,col2 combination to split data 
```
## Descriptive Table 
```{r}

library(table1)
library(dplyr)
library(knitr)

#check demographic variable for each subject (not for each observaion)
dat1_subject=dat1[!duplicated(dat1$id), ]
table1::table1(~age+biol_sex+PTSDorNot+ethnicity_id+bmi+education+marital_status |Treatment_new,data=dat1_subject,format = "markdown")
```

```{r}
#check biomarker baseline level for active/sham
options(knitr.table.format = 'markdown')
dat1_baseline=dat1%>%filter(Timepoints==0)

table1::table1(~BDNF_Mean+IFNgamma_Mean+IL1beta_Mean+IL2_Mean+IL6_Mean+TNFa_Mean+Ghrelin_Mean+IL12p70_Mean+IL13_Mean+IL22_Mean+IL5_Mean+MIF_Mean |Treatment_new,data=dat1_baseline,format = "markdown")
```
## check missing patterns
```{r}
 library(visdat)
vis_miss(dat1)
```

##  Distribution plots for the biomarkers 
```{r}

histbio_sep=function(biom,data,col1,col2){
  dat=NULL
  for (j in na.omit(unique(data[[col1]]))){
    for (h in na.omit(unique(data[[col2]]))){
      gpname=paste0(j,h)      
      gp=data.frame(subset(data,(data[[col1]]==j&data[[col2]]==h))[[biom]],gpname=paste0(j,h))
      colnames(gp)=c('biom','gpname')
      #assign(gpname,gp)
      dat=rbind(dat,gp)
    }
  }
  #allname=apply(expand.grid(c('PTSD','non-PTSD'), c('active','sham')), 1, function(x) paste0(x[1], x[2])) 
  
  #aa=mget(allname)
  #aaa=do.call(qpcR:::cbind.na,aa)
  #pdf(paste0('group_hist_',biom,'.pdf'))
  library(ggplot2)
  par(mfrow=c(3,4))
  print(ggplot(dat, aes(biom,fill=gpname)) + 
          geom_histogram(alpha = 0.5, position = 'identity')+ggtitle(paste0('Histogram of ',biom)))
  #dev.off()
  
}



for( i in biom){
  histbio_sep(i,dat1,col1 = 'Group',col2='Treatment_new')
}

```

## The Frequency table group by ID 


```{r}


#how many id have 1,2... obs 
dat1%>% group_by(`id`)%>% summarise(Freq=n())%>%arrange(Freq)%>% group_by(Freq)%>% summarise(number_id=n())
# the obs for each subject
dat1%>% group_by(`id`)%>% summarise(Obs=n())%>%arrange(Obs)
```

```{r eval=FALSE, include=FALSE}
library(dplyr)
#table 2 


#table 1
freqtable=function(data,batch,col1,col2){
data$ID_Point=apply(data[ ,c(col1,col2) ] , 1 , paste , collapse = "_")
table1=data %>% group_by(`ID_Point`)%>%
  summarise(distinct= n_distinct(UQ(as.name(batch))))%>%group_by(distinct)%>% summarise(Freq=n())
table2=data %>%
  group_by(UQ(as.name(batch))) %>%
  summarise(distinct= n_distinct(UQ(as.name(col1))))#id
freq1 =data%>% group_by(UQ(as.name(col1)),UQ(as.name(batch))) %>% summarise(Freq=n())
freq2=data %>% group_by(UQ(as.name(batch)),UQ(as.name(col1)))%>% summarise(Freq=n())

#write.csv(table1,file=paste0(batch,'_freqtable1.csv'))
#write.csv(table2,file=paste0(batch,'_freqtable2.csv'))
#write.csv(freq1,file=paste0(batch,'_freq1.csv'))
#write.csv(freq2,file=paste0(batch,'_freq2.csv'))
return(list(table1,table2,freq1,freq2))
}
freqtable(dat1,'Batch_A',col1 = 'id', col2 = "Timepoints")
freqtable(dat1,'Batch_S',col1 = 'id', col2 = "Timepoints")


```
##   Create Regression Model 
###  Model 1 

```{r include=FALSE}
#step3 mixed model

library(dplyr)
dat1_trau=dat1 %>% mutate(Trauma =ifelse(Timepoints==4,1,
               ifelse(Timepoints==10, 1,
               ifelse(Timepoints==14, 1,0))))


mixmodel1=function(biom,data){
library(lmerTest)
var.baseline=paste0('baseline_',biom)
new=data%>%group_by(id)%>%mutate((!!var.baseline):=UQ(as.name(biom))[1])
new_data=subset(data.frame(new),Timepoints>0)

#do we need to remove biom when time=0
fit=lmer(new_data[[biom]]~age+factor(biol_sex)+Treatment+Trauma+PTSDorNot+Treatment:PTSDorNot+new_data[[var.baseline]]+(1|id),data=new_data)
library(sjPlot)
tab_model(fit)
p=summary(fit)
return(list('p'=p,'fit'=fit))
}


pvec=list()
 fitvec=list()
for( i in biom){
  p=mixmodel1(i,dat1_trau)$p
  fit=mixmodel1(i,dat1_trau)$fit
  pvec[[i]]=p
  fitvec[[i]]=fit
  par(mfrow=c(1,2))

qqnorm(resid(fit),main=i)
qqline(resid(fit))
hist(resid(fit),main=i)
print(plot(fit,main=i))

}


```
###  Model 1  Paremeter estimates :regress on treatment,group(PTSD),trauma,baseline biomarker+random subject intercept
```{r}
library(sjPlot)

tab_model(fitvec,dv.labels=names(fitvec),show.se = TRUE)
#interaction not significant ,remove it 
```


###  Model 1 :remove interaction 

```{r echo=FALSE}
#step3 mixed model

mixmodel1=function(biom,data){
library(lmerTest)
var.baseline=paste0('baseline_',biom)
new=data%>%group_by(id)%>%mutate((!!var.baseline):=UQ(as.name(biom))[1])
new_data=subset(data.frame(new),Timepoints>0)

#do we need to remove biom when time=0
fit=lmer(new_data[[biom]]~age+factor(biol_sex)+Treatment+Trauma+PTSDorNot+new_data[[var.baseline]]+(1|id),data=new_data)
library(sjPlot)
tab_model(fit)
p=summary(fit)
return(list('p'=p,'fit'=fit))
}


pvec=list()
 fitvec=list()
for( i in biom){
  p=mixmodel1(i,dat1_trau)$p
  fit=mixmodel1(i,dat1_trau)$fit
  pvec[[i]]=p
  fitvec[[i]]=fit
 

}


```
```{r}
library(sjPlot)

tab_model(fitvec,dv.labels=names(fitvec),show.se = TRUE)

```

###  Model 2 Paremeter estimates: regress on group(PTSD), trauma script, VNS ,baseline biomarker+random subject intercept

```{r include=FALSE}
# add VNS and Trauma variable for each observation 
# base on plasma data existing timepoints only 
# 0  4  6  8 10 12 14 15 18 19 22 NA
library(dplyr)
dat1_vns=dat1 %>%mutate(VNS=ifelse(Timepoints==4,1,
               ifelse((Timepoints/2)%%2==1,1,0)))%>% mutate(Trauma =ifelse(Timepoints==4,1,
               ifelse(Timepoints==10, 1,
               ifelse(Timepoints==14, 1,0))))

#step3 mixed model

mixmodel2=function(biom,data){
library(lmerTest)
var.baseline=paste0('baseline_',biom)
new=data%>%group_by(id)%>%mutate((!!var.baseline):=UQ(as.name(biom))[1])
new_data=subset(data.frame(new),Timepoints>0)

#do we need to remove biom when time=0
fit=lmer(new_data[[biom]]~age+factor(biol_sex)+Trauma+VNS+PTSDorNot+new_data[[var.baseline]]+(1|id),data=new_data)
library(sjPlot)
tab_model(fit)
p=summary(fit)
return(list('p'=p,'fit'=fit))
}

pvec2=list()
 fitvec2=list()
for( i in biom){
  p=mixmodel2(i,dat1_vns)$p
  fit=mixmodel2(i,dat1_vns)$fit
  pvec2[[i]]=p
  fitvec2[[i]]=fit
   par(mfrow=c(1,2))

qqnorm(resid(fit),main=i)
qqline(resid(fit))
hist(resid(fit),main=i)
print(plot(fit,main=i))
}


```
```{r}
library(sjPlot)

tab_model(fitvec2,dv.labels=names(fitvec2),show.se = TRUE)
```

###  Model 2 :add VNS:PTSD interaction 
```{r echo=FALSE}
#step3 mixed model

mixmodel2=function(biom,data){
library(lmerTest)
var.baseline=paste0('baseline_',biom)
new=data%>%group_by(id)%>%mutate((!!var.baseline):=UQ(as.name(biom))[1])
new_data=subset(data.frame(new),Timepoints>0)

#do we need to remove biom when time=0
fit=lmer(new_data[[biom]]~age+factor(biol_sex)+Trauma+VNS+PTSDorNot+VNS:PTSDorNot+new_data[[var.baseline]]+(1|id),data=new_data)
library(sjPlot)
tab_model(fit)
p=summary(fit)
return(list('p'=p,'fit'=fit))
}

pvec2=list()
 fitvec2=list()
for( i in biom){
  p=mixmodel2(i,dat1_vns)$p
  fit=mixmodel2(i,dat1_vns)$fit
  pvec2[[i]]=p
  fitvec2[[i]]=fit
   par(mfrow=c(1,2))


}

```
```{r}
library(sjPlot)

tab_model(fitvec2,dv.labels=names(fitvec2),show.se = TRUE)
```

